---
title: "Figure 7"
author: "Lisa Holsten"
date: 2023-09-04
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Packages
```{r}
library(huva)
library(huva.db)
library(useful)
library(limma)
library(fgsea)
library(dplyr)
library(ggrepel)
library(pheatmap)
library(Rmisc)
library(ggplot2)
library(reshape2)
library(ggpubr)
library(patchwork)
```

## Functions
```{r}
source("run_huva_experiment_reverse.R")

source("plot_binned_gene_modified.R")

scaleColors <- function(data = input_scale, # data to use
                        maxvalue = NULL # value at which the color is fully red / blue
                        ){
  if(is.null(maxvalue)){
    maxvalue <- floor(min(abs(min(data)), max(data)))
  }
  if(max(data) > abs(min(data))){
    if(ceiling(max(data)) == maxvalue){
      myBreaks <- c(floor(-max(data)), seq(-maxvalue+0.2, maxvalue-0.2, 0.2),  ceiling(max(data)))
    } else{
      myBreaks <- c(floor(-max(data)), seq(-maxvalue, maxvalue, 0.2),  ceiling(max(data)))
    }
    paletteLength <- length(myBreaks)
    myColor <- colorRampPalette(c("blue", "white", "red"))(paletteLength)
  } else {
    if(-floor(min(data)) == maxvalue){
      myBreaks <- c(floor(min(data)), seq(-maxvalue+0.2, maxvalue-0.2, 0.2),  ceiling(min(data)))
    } else{
      myBreaks <- c(floor(min(data)), seq(-maxvalue, maxvalue, 0.2),  ceiling(abs(min(data))))
    }
    paletteLength <- length(myBreaks)
    myColor <- colorRampPalette(c("blue", "white", "red"))(paletteLength)
  }
 return(list(breaks = myBreaks, color = myColor))
}
```

```{r}
set.seed(42)
```


# ---

# 1. E2F1

### Expression of E2F1 in selected dataset

```{r, fig.height=4, fig.width=5}
gene_name <- "E2F1"
gene_overview <- gene_exam(huva_dataset = huva.db, gene = gene_name)
expr_exam <- get_expr_exam(huva_expression = gene_overview, study = "FG500", dataset = "FG500_whole_blood")

ggplot(expr_exam, aes(x=expression)) +
  geom_histogram(color="black", bins = 50, fill= "white") + 
  theme_bw() +
  geom_vline(xintercept =quantile(expr_exam$expression, 0.1), linetype = "dashed") + 
  geom_vline(xintercept =quantile(expr_exam$expression, 0.9), linetype = "dashed") + 
  annotate("rect", xmin=-Inf, xmax=quantile(expr_exam$expression, 0.1), ymin=0, ymax=Inf, alpha=0.2, fill="#005596") +  
  annotate("rect", xmin=quantile(expr_exam$expression, 0.9), xmax=Inf, ymin=0, ymax=Inf, alpha=0.2, fill="#6b1414") + 
  xlab("E2F1 vst expression") + ylab("Number of donors") 

ggsave(filename = "plots/Fig_7B_huva_E2F1_histogram.pdf", device = pdf, dpi = "print", height = 3, width = 3)
```


### Binning by E2F1 expression into low and high group
```{r}
binned_dataset <- run_huva_experiment_rev(data = huva.db, 
                                      gene = gene_name, 
                                      quantiles = 0.1, 
                                      gs_list = hallmarks_V7.2,
                                      summ = T, 
                                      datasets_list = NULL, 
                                      adjust.method = "none")
```

```{r, fig.height=4, fig.width=9}
plot_binned <- plot_binned_gene_mod(goi = c("E2F1"), huva_experiment = binned_dataset)
p0 <- plot_binned$FG500_whole_blood + ylim(c(0, 15)) + ggtitle("")

plot_binned <- plot_binned_gene_mod(goi = c("MYB"), huva_experiment = binned_dataset)
p1 <- plot_binned$FG500_whole_blood + ylim(c(0, 15)) + ggtitle("")

plot_binned <- plot_binned_gene_mod(goi = c("STAT1"), huva_experiment = binned_dataset)
p2 <- plot_binned$FG500_whole_blood + ylim(c(0, 15)) + ggtitle("")

plot_binned <- plot_binned_gene_mod(goi = c("FLI1"), huva_experiment = binned_dataset)
p3 <- plot_binned$FG500_whole_blood + ylim(c(0, 15)) + ggtitle("")

p0 + p1 + p2 + p3 + plot_layout(ncol = 4, guides = "collect")

ggsave(filename = "plots/Fig_7C_E2F1_genes.pdf", device = pdf, dpi = "print", height = 4, width = 9)
```


### GSEA on ranked gene list

```{r}
gsea_huva <- get_gsea_huva(huva_exp = binned_dataset, study = "FG500", int_plot = F)
```

```{r, fig.height=3, fig.width=5}
tmp <- gsea_huva$FG500_whole_blood[abs(gsea_huva$FG500_whole_blood$NES)>2, ] %>% arrange(NES)
tmp$pathway <- stringr::str_remove_all(string = tmp$pathway, pattern = "HALLMARK_") %>% stringr::str_replace_all("_", " ")

ggplot(tmp, aes(x=forcats::fct_reorder(pathway, NES), y=NES)) +
  geom_bar(stat="identity") + 
  coord_flip() + 
  ylim(-3.5, 3.5) + 
  geom_hline(yintercept = 0) + 
  theme_bw() + 
  theme(panel.grid = element_blank()) + 
  theme(legend.position = "none") + 
  xlab("")
ggsave("plots/Fig_7D_E2F1enrich.pdf", device = pdf, dpi = "print", height = 3, width = 5)
```


# ---

# 2. STAT1

### Expression of STAT1 in selected dataset

```{r, fig.height=4, fig.width=5}
gene_name <- "STAT1"
gene_overview <- gene_exam(huva_dataset = huva.db, gene = gene_name)
expr_exam <- get_expr_exam(huva_expression = gene_overview, study = "FG500", dataset = "FG500_whole_blood")

ggplot(expr_exam, aes(x=expression)) +
  geom_histogram(color="black", bins = 50, fill= "white") + 
  theme_bw() +
  geom_vline(xintercept =quantile(expr_exam$expression, 0.1), linetype = "dashed") + 
  geom_vline(xintercept =quantile(expr_exam$expression, 0.9), linetype = "dashed") + 
  annotate("rect", xmin=-Inf, xmax=quantile(expr_exam$expression, 0.1), ymin=0, ymax=Inf, alpha=0.2, fill="#005596") +  
  annotate("rect", xmin=quantile(expr_exam$expression, 0.9), xmax=Inf, ymin=0, ymax=Inf, alpha=0.2, fill="#6b1414") + 
  xlab("STAT1 vst expression") + ylab("Number of donors") 

ggsave(filename = "plots/Fig_7H_STAT1_histogram.pdf", device = pdf, dpi = "print", height = 3, width = 3)
```


### Binning by STAT1 expression into low and high group

```{r}
source("run_huva_experiment_reverse.R")
binned_dataset <- run_huva_experiment_rev(data = huva.db, 
                                          gene = gene_name, 
                                          quantiles = 0.1, 
                                          gs_list = hallmarks_V7.2,
                                          summ = T, 
                                          datasets_list = NULL, 
                                          adjust.method = "none")
```

```{r, fig.height=4, fig.width=9}
plot_binned <- plot_binned_gene_mod(goi = c("E2F1"), huva_experiment = binned_dataset)
p0 <- plot_binned$FG500_whole_blood + ylim(c(0, 15)) + ggtitle("")

plot_binned <- plot_binned_gene_mod(goi = c("MYB"), huva_experiment = binned_dataset)
p1 <- plot_binned$FG500_whole_blood + ylim(c(0, 15)) + ggtitle("")

plot_binned <- plot_binned_gene_mod(goi = c("STAT1"), huva_experiment = binned_dataset)
p2 <- plot_binned$FG500_whole_blood + ylim(c(0, 15)) + ggtitle("")

plot_binned <- plot_binned_gene_mod(goi = c("FLI1"), huva_experiment = binned_dataset)
p3 <- plot_binned$FG500_whole_blood + ylim(c(0, 15)) + ggtitle("")

p0 + p1 + p2 + p3 + plot_layout(ncol = 4, guides = "collect")

ggsave(filename = "plots/Fig_7I_STAT1_genes.pdf", device = pdf, dpi = "print", height = 4, width = 9)
```

### GSEA on ranked gene list

```{r}
gsea_huva <- get_gsea_huva(huva_exp = binned_dataset, study = "FG500", int_plot = F)
```

```{r, fig.height=3, fig.width=5}
tmp <- gsea_huva$FG500_whole_blood[abs(gsea_huva$FG500_whole_blood$NES)>2, ] %>% arrange(NES)
tmp$pathway <- stringr::str_remove_all(string = tmp$pathway, pattern = "HALLMARK_") %>% stringr::str_replace_all("_", " ")

ggplot(tmp, aes(x=forcats::fct_reorder(pathway, NES), y=NES)) +
  geom_bar(stat="identity") + 
  coord_flip() + 
  ylim(-3.5, 3.5) + 
  geom_hline(yintercept = 0) + 
  theme_bw() + 
  theme(panel.grid = element_blank()) + 
  theme(legend.position = "none") + 
  xlab("")
ggsave("plots/Fig_7J_STAT1_enrich.pdf", device = pdf, dpi = "print", height = 3, width = 4)
```


# ---

# 3. FLI1

### Expression of FLI1 in selected dataset

```{r, fig.height=4, fig.width=5}
gene_name <- "FLI1"
gene_overview <- gene_exam(huva_dataset = huva.db, gene = gene_name)
expr_exam <- get_expr_exam(huva_expression = gene_overview, study = "FG500", dataset = "FG500_whole_blood")

ggplot(expr_exam, aes(x=expression)) +
  geom_histogram(color="black", bins = 50, fill= "white") + 
  theme_bw() +
  geom_vline(xintercept =quantile(expr_exam$expression, 0.1), linetype = "dashed") + 
  geom_vline(xintercept =quantile(expr_exam$expression, 0.9), linetype = "dashed") + 
  annotate("rect", xmin=-Inf, xmax=quantile(expr_exam$expression, 0.1), ymin=0, ymax=Inf, alpha=0.2, fill="#005596") +  
  annotate("rect", xmin=quantile(expr_exam$expression, 0.9), xmax=Inf, ymin=0, ymax=Inf, alpha=0.2, fill="#6b1414") + 
  xlab("FLI1 vst expression") + ylab("Number of donors") 

ggsave(filename = "plots/Fig_7K_huva_FLI1_histogram.pdf", device = pdf, dpi = "print", height = 4, width = 5)
```


### Binning by FLI1 expression into low and high group

```{r}
binned_dataset <- run_huva_experiment_rev(data = huva.db, 
                                      gene = gene_name, 
                                      quantiles = 0.1, 
                                      gs_list = hallmarks_V7.2,
                                      summ = T, 
                                      datasets_list = NULL, 
                                      adjust.method = "none")
```

```{r, fig.height=4, fig.width=9}
plot_binned <- plot_binned_gene_mod(goi = c("E2F1"), huva_experiment = binned_dataset)
p0 <- plot_binned$FG500_whole_blood + ylim(c(0, 15)) + ggtitle("")

plot_binned <- plot_binned_gene_mod(goi = c("MYB"), huva_experiment = binned_dataset)
p1 <- plot_binned$FG500_whole_blood + ylim(c(0, 15)) + ggtitle("")

plot_binned <- plot_binned_gene_mod(goi = c("STAT1"), huva_experiment = binned_dataset)
p2 <- plot_binned$FG500_whole_blood + ylim(c(0, 15)) + ggtitle("")

plot_binned <- plot_binned_gene_mod(goi = c("FLI1"), huva_experiment = binned_dataset)
p3 <- plot_binned$FG500_whole_blood + ylim(c(0, 15)) + ggtitle("")

p0 + p1 + p2 + p3 + plot_layout(ncol = 4, guides = "collect")

ggsave(filename = "plots/Fig_7L_FLI1_genes.pdf", device = pdf, dpi = "print", height = 4, width = 9)
```


### GSEA on ranked gene list

```{r}
gsea_huva <- get_gsea_huva(huva_exp = binned_dataset, study = "FG500", int_plot = F)
```

```{r, fig.height=3, fig.width=5}
tmp <- gsea_huva$FG500_whole_blood[abs(gsea_huva$FG500_whole_blood$NES)>2, ] %>% arrange(NES)
tmp$pathway <- stringr::str_remove_all(string = tmp$pathway, pattern = "HALLMARK_") %>% stringr::str_replace_all("_", " ")

ggplot(tmp, aes(x=forcats::fct_reorder(pathway, NES), y=NES)) +
  geom_bar(stat="identity") + 
  coord_flip() + 
  ylim(-3.5, 3.5) + 
  geom_hline(yintercept = 0) + 
  theme_bw() + 
  theme(panel.grid = element_blank()) + 
  theme(legend.position = "none") + 
  xlab("")
ggsave("plots/Fig_7M_FLI1_enrich.pdf", device = pdf, dpi = "print", height = 3, width = 5)
```

# ---

# 4. MYB

### Expression of MYB in selected dataset

```{r, fig.height=4, fig.width=5}
gene_name <- "MYB"
gene_overview <- gene_exam(huva_dataset = huva.db, gene = gene_name)
expr_exam <- get_expr_exam(huva_expression = gene_overview, study = "FG500", dataset = "FG500_whole_blood")

ggplot(expr_exam, aes(x=expression)) +
  geom_histogram(color="black", bins = 50, fill= "white") + 
  theme_bw() +
  geom_vline(xintercept =quantile(expr_exam$expression, 0.1), linetype = "dashed") + 
  geom_vline(xintercept =quantile(expr_exam$expression, 0.9), linetype = "dashed") + 
  annotate("rect", xmin=-Inf, xmax=quantile(expr_exam$expression, 0.1), ymin=0, ymax=Inf, alpha=0.2, fill="#005596") +  
  annotate("rect", xmin=quantile(expr_exam$expression, 0.9), xmax=Inf, ymin=0, ymax=Inf, alpha=0.2, fill="#6b1414") + 
  xlab("MYB vst expression") + ylab("Number of donors") 

ggsave(filename = "plots/Fig_7E_huva_MYB_histogram.pdf", device = pdf, dpi = "print", height = 3, width = 3)
```


### Binning by MYB expression into low and high group

```{r}
source("run_huva_experiment_reverse.R")
binned_dataset <- run_huva_experiment_rev(data = huva.db, 
                                          gene = gene_name, 
                                          quantiles = 0.1, 
                                          gs_list = hallmarks_V7.2,
                                          summ = T, 
                                          datasets_list = NULL, 
                                          adjust.method = "none")
```

```{r, fig.height=4, fig.width=9}
plot_binned <- plot_binned_gene_mod(goi = c("E2F1"), huva_experiment = binned_dataset)
p0 <- plot_binned$FG500_whole_blood + ylim(c(0, 15)) + ggtitle("")

plot_binned <- plot_binned_gene_mod(goi = c("MYB"), huva_experiment = binned_dataset)
p1 <- plot_binned$FG500_whole_blood + ylim(c(0, 15)) + ggtitle("")

plot_binned <- plot_binned_gene_mod(goi = c("STAT1"), huva_experiment = binned_dataset)
p2 <- plot_binned$FG500_whole_blood + ylim(c(0, 15)) + ggtitle("")

plot_binned <- plot_binned_gene_mod(goi = c("FLI1"), huva_experiment = binned_dataset)
p3 <- plot_binned$FG500_whole_blood + ylim(c(0, 15)) + ggtitle("")

p0 + p1 + p2 + p3 + plot_layout(ncol = 4, guides = "collect")

ggsave(filename = "plots/Fig_7F_MYB_genes.pdf", device = pdf, dpi = "print", height = 4, width = 9)
```


### GSEA on ranked gene list

```{r}
gsea_huva <- get_gsea_huva(huva_exp = binned_dataset, study = "FG500", int_plot = F)
```

```{r, fig.height=3, fig.width=5}
tmp <- gsea_huva$FG500_whole_blood[abs(gsea_huva$FG500_whole_blood$NES)>2, ] %>% arrange(NES)
tmp$pathway <- stringr::str_remove_all(string = tmp$pathway, pattern = "HALLMARK_") %>% stringr::str_replace_all("_", " ")

ggplot(tmp, aes(x=forcats::fct_reorder(pathway, NES), y=NES)) +
  geom_bar(stat="identity") + 
  coord_flip() + 
  ylim(-3.5, 3.5) + 
  geom_hline(yintercept = 0) + 
  theme_bw() + 
  theme(panel.grid = element_blank()) + 
  theme(legend.position = "none") + 
  xlab("")
ggsave("plots/Fig_7G_MYB_enrich.pdf", device = pdf, dpi = "print", height = 3, width = 5)
```
