---
title: "Figure 6 & Figure S7"
author: "Lisa Holsten"
date: '2023-09-04'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Packages

```{r}
BiocManager::install("ggnetwork", update = FALSE)
library(ggnetwork)
```

```{r}
library(DESeq2)
library(sva)
library(limma)
library(ggplot2)
library(ggrepel)
library(tidyverse)
library(UpSetR)
library(dplyr)
library(tidyr)

library(ggpubr)
library(pheatmap)
library(ComplexHeatmap)
library(circlize)
library(RcisTarget)
library(stringr)
library(rstatix)
```


## Color scheme

```{r}
cols_module <- c("gold", "lightgreen", "maroon", "plum", "seagreen", "steelblue", "turquoise", "wheat" )
names(cols_module) <- cols_module

col_group <- c("#FC6B9D", "#323232")
names(col_group) <- c("NEO", "AD")

col_treatment <- c("#d4d6d8", "#82898e")
names(col_treatment) <- c("ctl", "LPS")
```


## Import
```{r}
sample_table <- read.table(file = "tables/sample_table.txt", 
                           sep = ";", 
                           stringsAsFactors = F)
sample_table$condition <- factor(sample_table$condition, levels = c("NEO_ctl", "AD_ctl", "NEO_LPS", "AD_LPS"))
```

```{r}
counts_bc <- readRDS("hCoCena_objects/counts_batchcorr_log.rds")
hcobject <- readRDS("hCoCena_objects/hcobject.rds")
gtc <- readRDS("hCoCena_objects/gtc.rds")
DE_results.df <- read.csv("tables/DEGs.csv",  sep = "\t", dec = ",")
```


# ---


# 1. Identifiy modules of interest based on differences in mean expression between groups at baseline

```{r}
# Calculate mean expression per sample and module & filter for control condition
mean.module <- counts_bc %>% mutate("gene" = rownames(counts_bc)) %>% left_join(gtc) %>% 
  tidyr::pivot_longer(cols = 1:12, names_to = "sample", values_to = "expression") %>% na.omit() %>% 
  group_by(color, sample) %>% summarize(mean_sample = mean(expression)) %>%
  left_join(sample_table, by = c("sample" = "ID")) 

## Calculate statistics
stats.df <- compare_means(formula = mean_sample ~ condition, 
                          data = mean.module, 
                          method = "t.test", 
                          paired = F,  
                          group.by = "color",
                          p.adjust.method = "bonferroni") 
stats.df$comparison <- paste0(stats.df$group1, "_", stats.df$group2)


stats.baseline <- stats.df[stats.df$comparison == "NEO_ctl_AD_ctl", ]
stats.baseline$p.adj <- round(p.adjust(p = stats.baseline$p, method = "bonferroni"), digits = 4)
```


### Heatmap (Fig 6A)

```{r}
# Matrix with mean expression per sample & gene name in last name ("Gene)
counts_scaled <- counts_bc[, colnames(counts_bc) %in% sample_table[sample_table$treatment == "ctl", ]$ID]

if (identical(colnames(counts_scaled), sample_table$ID[1:6])){
  colnames(counts_scaled) <- sample_table$donor[1:6]
} else {print("Reorder")}

counts_scaled <- t(scale(t(counts_scaled))) %>% as.data.frame() %>% mutate("Gene" = rownames(counts_scaled))
```

```{r, fig.height=5, fig.width=8}
source("hCoCena_objects/plot_cluster_heatmap_mean.R")

hcobject$global_settings$voi <- "donor"
plot_cluster_heatmap_mean(matrix_groupmean = counts_scaled,
                          stats = stats.baseline,
                          col_order = colnames(counts_scaled)[1:6],
                          row_order = c("maroon", "gold", "steelblue", "plum", "turquoise", "wheat", "seagreen", "lightgreen"), 
                          cluster_columns = F,
                          cluster_rows = F ,
                          file_name = "Fig_6A_ModuleHeatmap.pdf", 
                          return_HM = T)
```


### Boxplot (Fig S7A)

```{r, fig.height=6, fig.width=8}
mean.module$condition <- factor(mean.module$condition, levels = c("NEO_ctl", "AD_ctl", "NEO_LPS", "AD_LPS"))
mean.module$color <- factor(mean.module$color, levels = c("maroon", "gold", "steelblue", "plum", "turquoise", "wheat", "seagreen", "lightgreen"))

stats <- stats.df[stats.df$comparison %in% c("NEO_ctl_AD_ctl", "NEO_LPS_AD_LPS"), ]
stats$p.adj <- round(p.adjust(p = stats$p, method = "bonferroni"), digits = 4)

stat.test <- stats %>%
  mutate(y.position = 9) %>% 
  dplyr::select(-.y.) %>%
  mutate(p.adj.signif = symnum(p.adj, corr = FALSE, na = FALSE, 
                               cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), 
                               symbols = c("****", "***", "**", "*", "ns")))
stat.test$color <- factor(stat.test$color, levels = c("maroon", "gold", "steelblue", "plum", "turquoise", "wheat", "seagreen", "lightgreen"))

ggplot(mean.module, aes(x=condition, y=mean_sample)) + 
  geom_boxplot(alpha = 0.5) +
  facet_wrap(~color, ncol = 4) + 
  ylim(c(0, 11)) + 
  stat_pvalue_manual(data = stat.test, label = "p.adj.signif") + 
  geom_jitter(size=1) +
  facet_wrap(~color, ncol = 4) +
  ylim(c(0, 11)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
  xlab("") + ylab("mean expression") 

ggsave(filename = "plots/Fig_S7A_MeanBoxplot.pdf", device = cairo_pdf, dpi = "print", height = 4, width = 6)
```


# ---


# 2. Transcription factor-binding site prediction analysis using on RCisTarget

Based on genes within modules of interest as determined in A
```{r}
# gene lists
genes_maroon     <- gtc[gtc$color == "maroon", ]$gene
genes_lightgreen <- gtc[gtc$color == "lightgreen", ]$gene
genes_steelblue  <- gtc[gtc$color == "steelblue", ]$gene

input_list <- list("module_maroon"     = genes_maroon,
                   "module_lightgreen" = genes_lightgreen,
                   "module_steelblue"  = genes_steelblue)

# Search space: 10k bp around TSS - HUMAN
motifRankings <- importRankings("~/downstream_references/downstream_references/rcistarget/hg38__refseq-r80__10kb_up_and_down_tss.mc9nr_.feather")

# Load the annotation to human transcription factors
data(motifAnnotations_hgnc)

# Calculate enrichment
motifEnrichmentTable <- cisTarget(input_list,
                                  motifRankings,
                                  motifAnnot=motifAnnotations_hgnc)

# Transform
annotatedTfs <- lapply(split(motifEnrichmentTable$TF_highConf, motifEnrichmentTable$geneSet), function(x) {
  genes <- gsub(" \\(.*\\). ", "; ", x, fixed=FALSE)
  genesSplit <- unique(unlist(strsplit(genes, "; ")))
  return(genesSplit)
  })
```


### Heatmap of candidate regulators (Fig 6B)

```{r}
candidate_regulators <- annotatedTfs$module_steelblue[annotatedTfs$module_steelblue %in% 
                                                        c(annotatedTfs$module_lightgreen, annotatedTfs$module_maroon)]

DEGs_baseline <- DE_results.df[DE_results.df$comparison == "NEO_ctl vs AD_ctl", ]

matrix <- counts_bc[rownames(counts_bc) %in% candidate_regulators, colnames(counts_bc) %in% sample_table[sample_table$treatment == "ctl", ]$ID] 
if(identical(colnames(matrix), sample_table[sample_table$treatment == "ctl", ]$ID)){
  colnames(matrix) <- sample_table[sample_table$treatment == "ctl", ]$donor
} else {print("Re-order")}
```

```{r}
# Row annotation
anno_row_df <- data.frame(gene = rownames(matrix)) %>% 
  left_join(DEGs_baseline[, c("SYMBOL", "regulation")], by = c("gene" = "SYMBOL")) %>%
  left_join(gtc) %>%
  column_to_rownames("gene")
colnames(anno_row_df) <- c("DE_NEOctl_vs_ADctl", "CoCena_module")

anno_row_df$TFBS_maroon     <- ifelse(rownames(anno_row_df) %in% annotatedTfs$module_maroon, "yes", "no")
anno_row_df$TFBS_lightgreen <- ifelse(rownames(anno_row_df) %in% annotatedTfs$module_lightgreen, "yes", "no")
anno_row_df$TFBS_steelblue  <- ifelse(rownames(anno_row_df) %in% annotatedTfs$module_steelblue, "yes", "no")

anno_row_df$CoCena_module <- factor(anno_row_df$CoCena_module, 
                                    levels = c("maroon", "gold", "steelblue", "plum", "turquoise", "wheat", "seagreen", "lightgreen"))

# Column annotation
anno_col_df <- sample_table[sample_table$donor %in% colnames(matrix), ]  %>% 
  dplyr::select(donor, group) %>% unique() %>% remove_rownames()  %>% column_to_rownames("donor")


# Colors
cols_DEA <- c("dodgerblue3", "firebrick4", "lightgrey")
names(cols_DEA) <- c("down", "up", "n.s.")

cols_module <- unique(gtc$color)
names(cols_module) <- unique(gtc$color)

cols_TFBS_m <- c("maroon", "lightgrey")
names(cols_TFBS_m) <- c("yes", "no")

cols_TFBS_l <- c("lightgreen", "lightgrey")
names(cols_TFBS_l) <- c("yes", "no")

cols_TFBS_s <- c("steelblue", "lightgrey")
names(cols_TFBS_s) <- c("yes", "no")

anno_colors <- list(group = col_group,
                    CoCena_module = cols_module, 
                    DE_NEOctl_vs_ADctl = cols_DEA,
                    TFBS_maroon = cols_TFBS_m,
                    TFBS_lightgreen = cols_TFBS_l, 
                    TFBS_steelblue = cols_TFBS_s)
```

```{r, fig.height=15, fig.width=7}
pheatmap::pheatmap(mat = matrix, 
         cluster_rows = T, 
         cluster_cols = F,
         scale = "row", 
         cutree_rows = 2,
         annotation_col = anno_col_df,
         annotation_colors = anno_colors,
         annotation_row = anno_row_df, 
         na_col = "lightgrey",
         cellwidth = 15, 
         cellheight = 15, 
         treeheight_row = 10, 
         drop_levels = T,
         main = "Candidate regulators", 
         filename = "plots/Fig_6B_Heatmap_candidate_regulators.pdf", 
         width = 6, height = 7)
```


### Heatmap of E2F1 target genes (Fig S7B)

```{r}
# Get all potential E2F1 target genes
E2F1_df <- motifEnrichmentTable[str_detect(string = motifEnrichmentTable$TF_highConf, pattern = "E2F1", negate = F), ]

E2F1_targets <- lapply(split(x = E2F1_df$enrichedGenes, f = E2F1_df$geneSet), function(x) {
  genes <- gsub(" \\(.*\\). ", "; ", x, fixed=FALSE)
  genesSplit <- unique(unlist(strsplit(genes, ";")))
  return(genesSplit)
  }) %>% unlist() %>% unique()

# Calculate mean expression per condition
mean.group <- counts_bc %>% mutate("gene" = rownames(counts_bc)) %>% 
  pivot_longer(cols = 1:12, names_to = "sample", values_to = "expression") %>% 
  left_join(sample_table[, c("ID", "condition")], by = c("sample" = "ID")) %>% 
  group_by(gene, condition) %>% summarize(mean_group = mean(expression)) %>%
  pivot_wider(id_cols = "gene", names_from = "condition", values_from = mean_group) %>%
  column_to_rownames("gene")

# Matrix
matrix <- mean.group[rownames(mean.group) %in% E2F1_targets, ]
matrix <- matrix[, c("NEO_ctl", "AD_ctl", "NEO_LPS", "AD_LPS")]
```

```{r}
# Column annotation
anno_col_df <- sample_table[sample_table$condition %in% colnames(matrix), ]  %>% dplyr::select(condition, group, treatment) %>% unique() %>%
  remove_rownames() %>%  column_to_rownames("condition")

# Row annotation
DEGenes_wide <- DE_results.df %>% pivot_wider(id_cols = "SYMBOL", names_from = "comparison", values_from = "regulation")

anno_row_df <- data.frame(gene = rownames(matrix)) %>% 
  left_join(DEGenes_wide, by = c("gene" = "SYMBOL")) %>%
  left_join(gtc) %>%
  column_to_rownames("gene")
colnames(anno_row_df) <- c("NEO_ctlvsAD_ctl", "NEO_LPSvsAD_LPS", "NEO_LPS_vs_NEO_ctl", "AD_LPSvsAD_ctl", "CoCena_module")

# Colors
anno_colors <- list(group              = col_group,
                    treatment          = col_treatment,
                    CoCena_module      = cols_module[names(cols_module) %in% c("maroon", "steelblue", "lightgreen")], 
                    NEO_ctlvsAD_ctl    = cols_DEA,
                    NEO_LPSvsAD_LPS    = cols_DEA,
                    NEO_LPS_vs_NEO_ctl = cols_DEA, 
                    AD_LPSvsAD_ctl     = cols_DEA)
```


```{r, fig.height=13, fig.width=5}
pheatmap::pheatmap(mat = matrix, 
         cluster_rows = T, 
         cluster_cols = F,
         scale = "row", 
         show_rownames = F,
         annotation_col = anno_col_df,
         annotation_colors = anno_colors,
         annotation_row = anno_row_df, 
         cellwidth = 15, 
         treeheight_row = 10,
         main = "E2F1 target genes", 
         filename = "plots/Fig_S7B_Heatmap_E2F1_targets.pdf",
         width = 6, height = 8)
```


### Network visualization of potential regulators and their targets

#### All (Fig 6D)
Network of all candidate regulators and their target TFs predicted to also regulate the modules of interest
```{r}
graph_df <- data.frame()

TFs_of_interest <- candidate_regulators
all_TFs <- c(annotatedTfs$module_lightgreen, annotatedTfs$module_maroon, annotatedTfs$module_steelblue)

# Get all target TFs of candidate regulators
for (i in 1:length(TFs_of_interest)){
  TF <- TFs_of_interest[i]
  tmp <- motifEnrichmentTable[str_detect(string = motifEnrichmentTable$TF_highConf, pattern = TF, negate = F), ]
  targets <- lapply(split(x = tmp$enrichedGenes, f = tmp$geneSet), function(x) {
    genes <- gsub(" \\(.*\\). ", "; ", x, fixed=FALSE)
    genesSplit <- unique(unlist(strsplit(genes, ";")))
    return(genesSplit)
    })
  target_genes <- targets %>% unlist() %>% unique()
  df <- data.frame("from" = rep(TF, length(target_genes)), "to" = target_genes)
  df <- df[!df$to == TF & df$to %in% all_TFs, ]
  
  graph_df <- rbind(graph_df, df)
}

# Generate network object
n <- network::network(graph_df, directed = T)
```

```{r, fig.height=7, fig.width=10}
n_targets <- graph_df %>% group_by(from) %>% summarize(n_targets = n())
n_regulators <- graph_df %>% group_by(to) %>% summarize(n_regulators = n())

set.seed(75)

df <- ggnetwork(n,  layout = "fruchtermanreingold", arrow.gap = 0.025) %>% 
  left_join(gtc, by = c("vertex.names" = "gene")) %>% 
  left_join(n_targets, by = c("vertex.names" = "from")) %>%
  left_join(n_regulators, by = c("vertex.names" = "to")) %>% 
  rowwise() %>% mutate(n_connections = sum(n_targets, n_regulators, na.rm = T))

# Include DE information
df <- left_join(df, DE_results.df[DE_results.df$comparison == "NEO_ctl vs AD_ctl", c("SYMBOL", "regulation")], by = c("vertex.names" = "SYMBOL"))

df$color <- ifelse(is.na(df$color), "not incl.", df$color)

cols_module_ext <-  c(unique(gtc$color), "grey")
names(cols_module_ext) <- c(unique(gtc$color), "not incl.")
```



```{r, fig.height=8, fig.width=12}
ggplot(df, aes(x = x, y = y, xend = xend, yend = yend)) +
  
  # Edges
  geom_edges(arrow = arrow(length = unit(6, "pt"), type = "open", angle=20), color = "grey50", size = 0.2) +
  
  # Nodes - DEA
  geom_nodes(aes(x, y, color = regulation, size = n_connections*1.5), shape = 21, stroke = 2, show.legend = T, fill = NA) + 
  scale_color_manual(values = cols_DEA) +
  ggnewscale::new_scale_color() + 
  
  # Nodes
  geom_nodes(aes(x, y, fill = color, size = n_connections), shape = 21, stroke = 1, color = "white") +
  scale_fill_manual(values = scales::alpha(cols_module_ext, 0.7)) +

  
  scale_size_continuous(breaks = c(1, 10, 20, 30),
                        labels = c("1", "10", "20","30"),
                        range = c(5, 15)) +
  
  # Node text
  geom_nodetext(aes(label = vertex.names), size = 5) +
  
  # General
  theme_blank()+ 
  ggtitle("")

ggsave(filename = "plots/Fig_6D_Network_regulators.pdf", device = cairo_pdf, dpi = "print", height = 8, width = 12)
```


#### Individual TFs (Fig S7C)

Network of E2F1 and its target TFs predicted to also regulate the modules of interest
```{r}
graph_df <- data.frame()

TFs_of_interest <- "E2F1"
all_TFs <- c(annotatedTfs$module_lightgreen, annotatedTfs$module_maroon, annotatedTfs$module_steelblue)

# Get all target TFs of candidate regulators
for (i in 1:length(TFs_of_interest)){
  TF <- TFs_of_interest[i]
  tmp <- motifEnrichmentTable[str_detect(string = motifEnrichmentTable$TF_highConf, pattern = TF, negate = F), ]
  targets <- lapply(split(x = tmp$enrichedGenes, f = tmp$geneSet), function(x) {
    genes <- gsub(" \\(.*\\). ", "; ", x, fixed=FALSE)
    genesSplit <- unique(unlist(strsplit(genes, ";")))
    return(genesSplit)
    })
  target_genes <- targets %>% unlist() %>% unique()
  df <- data.frame("from" = rep(TF, length(target_genes)), "to" = target_genes)
  df <- df[!df$to == TF & df$to %in% all_TFs, ]
  
  graph_df <- rbind(graph_df, df)
}

# Generate network object
n <- network::network(graph_df, directed = T)
```

```{r, fig.height=7, fig.width=10}
n_targets <- graph_df %>% group_by(from) %>% summarize(n_targets = n())
n_regulators <- graph_df %>% group_by(to) %>% summarize(n_regulators = n())

set.seed(10)
df <- ggnetwork(n,  layout = "fruchtermanreingold", arrow.gap = 0.05) %>% 
  left_join(gtc, by = c("vertex.names" = "gene"))
```

```{r, fig.height=5, fig.width=5}
ggplot(df, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(arrow = arrow(length = unit(6, "pt"), type = "open", angle=20), color = "grey50", size = 0.2) +
  geom_nodelabel(aes(label = vertex.names, fill = color), label.size = 0) +
  scale_fill_manual(values = cols_module) +
  theme_blank()+ 
  theme(legend.position = "none") + 
  ggtitle("")

ggsave(filename = "plots/Fig_S7C_Network_E2F1.pdf", device = cairo_pdf, dpi = "print", height = 5, width = 5)
```


Network of MYB and its target TFs predicted to also regulate the modules of interest
```{r}
graph_df <- data.frame()

TFs_of_interest <- "MYB"
all_TFs <- c(annotatedTfs$module_lightgreen, annotatedTfs$module_maroon, annotatedTfs$module_steelblue)

# Get all target TFs of candidate regulators
for (i in 1:length(TFs_of_interest)){
  TF <- TFs_of_interest[i]
  tmp <- motifEnrichmentTable[str_detect(string = motifEnrichmentTable$TF_highConf, pattern = TF, negate = F), ]
  targets <- lapply(split(x = tmp$enrichedGenes, f = tmp$geneSet), function(x) {
    genes <- gsub(" \\(.*\\). ", "; ", x, fixed=FALSE)
    genesSplit <- unique(unlist(strsplit(genes, ";")))
    return(genesSplit)
    })
  target_genes <- targets %>% unlist() %>% unique()
  df <- data.frame("from" = rep(TF, length(target_genes)), "to" = target_genes)
  df <- df[!df$to == TF & df$to %in% all_TFs, ]
  
  graph_df <- rbind(graph_df, df)
}

# Generate network object
n <- network::network(graph_df, directed = T)
```

```{r, fig.height=7, fig.width=10}
n_targets <- graph_df %>% group_by(from) %>% summarize(n_targets = n())
n_regulators <- graph_df %>% group_by(to) %>% summarize(n_regulators = n())

set.seed(10)
df <- ggnetwork(n,  layout = "fruchtermanreingold", arrow.gap = 0.05) %>% 
  left_join(gtc, by = c("vertex.names" = "gene"))
```

```{r, fig.height=5, fig.width=5}
ggplot(df, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(arrow = arrow(length = unit(6, "pt"), type = "open", angle=20), color = "grey50", size = 0.2) +
  geom_nodelabel(aes(label = vertex.names, fill = color), label.size = 0) +
  scale_fill_manual(values = cols_module) +
  theme_blank()+ 
  theme(legend.position = "none") + 
  ggtitle("")

ggsave(filename = "plots/Fig_S7C_Network_MYB.pdf", device = cairo_pdf, dpi = "print", height = 5, width = 5)
```


Network of STAT1 and its target TFs predicted to also regulate the modules of interest
```{r}
graph_df <- data.frame()

TFs_of_interest <- "STAT1"
all_TFs <- c(annotatedTfs$module_lightgreen, annotatedTfs$module_maroon, annotatedTfs$module_steelblue)

# Get all target TFs of candidate regulators
for (i in 1:length(TFs_of_interest)){
  TF <- TFs_of_interest[i]
  tmp <- motifEnrichmentTable[str_detect(string = motifEnrichmentTable$TF_highConf, pattern = TF, negate = F), ]
  targets <- lapply(split(x = tmp$enrichedGenes, f = tmp$geneSet), function(x) {
    genes <- gsub(" \\(.*\\). ", "; ", x, fixed=FALSE)
    genesSplit <- unique(unlist(strsplit(genes, ";")))
    return(genesSplit)
    })
  target_genes <- targets %>% unlist() %>% unique()
  df <- data.frame("from" = rep(TF, length(target_genes)), "to" = target_genes)
  df <- df[!df$to == TF & df$to %in% all_TFs, ]
  
  graph_df <- rbind(graph_df, df)
}

# Generate network object
n <- network::network(graph_df, directed = T)
```

```{r, fig.height=7, fig.width=10}
n_targets <- graph_df %>% group_by(from) %>% summarize(n_targets = n())
n_regulators <- graph_df %>% group_by(to) %>% summarize(n_regulators = n())

set.seed(10)
df <- ggnetwork(n,  layout = "fruchtermanreingold", arrow.gap = 0.05) %>% 
  left_join(gtc, by = c("vertex.names" = "gene"))
```

```{r, fig.height=5, fig.width=5}
ggplot(df, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(arrow = arrow(length = unit(6, "pt"), type = "open", angle=20), color = "grey50", size = 0.2) +
  geom_nodelabel(aes(label = vertex.names, fill = color), label.size = 0) +
  scale_fill_manual(values = cols_module) +
  theme_blank()+ 
  theme(legend.position = "none") + 
  ggtitle("")

ggsave(filename = "plots/Fig_S7C_Network_STAT1.pdf", device = cairo_pdf, dpi = "print", height = 5, width = 5)
```


Network of FLI1 and its target TFs predicted to also regulate the modules of interest
```{r}
graph_df <- data.frame()

TFs_of_interest <- "FLI1"
all_TFs <- c(annotatedTfs$module_lightgreen, annotatedTfs$module_maroon, annotatedTfs$module_steelblue)

# Get all target TFs of candidate regulators
for (i in 1:length(TFs_of_interest)){
  TF <- TFs_of_interest[i]
  tmp <- motifEnrichmentTable[str_detect(string = motifEnrichmentTable$TF_highConf, pattern = TF, negate = F), ]
  targets <- lapply(split(x = tmp$enrichedGenes, f = tmp$geneSet), function(x) {
    genes <- gsub(" \\(.*\\). ", "; ", x, fixed=FALSE)
    genesSplit <- unique(unlist(strsplit(genes, ";")))
    return(genesSplit)
    })
  target_genes <- targets %>% unlist() %>% unique()
  df <- data.frame("from" = rep(TF, length(target_genes)), "to" = target_genes)
  df <- df[!df$to == TF & df$to %in% all_TFs, ]
  
  graph_df <- rbind(graph_df, df)
}

# Generate network object
n <- network::network(graph_df, directed = T)
```

```{r, fig.height=7, fig.width=10}
n_targets <- graph_df %>% group_by(from) %>% summarize(n_targets = n())
n_regulators <- graph_df %>% group_by(to) %>% summarize(n_regulators = n())

set.seed(10)
df <- ggnetwork(n,  layout = "fruchtermanreingold", arrow.gap = 0.05) %>% 
  left_join(gtc, by = c("vertex.names" = "gene"))
```

```{r, fig.height=5, fig.width=5}
ggplot(df, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(arrow = arrow(length = unit(6, "pt"), type = "open", angle=20), color = "grey50", size = 0.2) +
  geom_nodelabel(aes(label = vertex.names, fill = color), label.size = 0) +
  scale_fill_manual(values = cols_module) +
  theme_blank()+ 
  theme(legend.position = "none") + 
  ggtitle("")

ggsave(filename = "plots/Fig_S7C_Network_FLI1.pdf", device = cairo_pdf, dpi = "print", height = 5, width = 5)
```


# ---


#3. Hub genes (hCoCena docker image; Fig 6C)

```{r}
hcocena::find_hubs(top = 15,
                   clusters = c("maroon", "steelblue", "lightgreen"),
                   save = T,
                   tree_layout = F,
                   TF_only = "TF",
                   Plot = T)
```



# ---

